### 2008-09-18 trunk/Papers/rmrs-08/introduction.tex
18/09/2008

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\newcite{bos:etal:2004} exploit {\sc ccg}'s elegant syntax-semantics
interface to produce a unique logical form for each derivation of a
statistical {\sc ccg} parser; \newcite{zettlemoyer:collins:2007} and
others learn string-to-logical-form mappings, exploting informative
features from shallow language processors; and
\newcite{copestake:2003} and \newcite{frank:2004} endow robust
language processors with semantic components.

@@

However, there are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields only 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a shallow
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferrable to
give the parser the option to to leave the decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation that neither over-determines nor under-determines the
semantic information revealed by the (shallow) syntactic analysis, and
to complete the semantic information later, using information
extraneous to the parser.

In this paper, we focus on one particular approach to representing
partial semantic information: Robust Minimal Recursion Semantics
(\rmrs, \cite{copestake:2003}).  \rmrs\ is a generalisation of \mrs\
\cite{copestake:etal:2005} that is designed to support
underspecification of scope and predicate-argument structure.  It is
an emerging standard representation language for partial semantic
information, and has been applied in several implemented systems.  For
instance, \newcite{copestake:2003} and \newcite{frank:2004} use it to
specify semantic components to shallow parsers ranging in depth from
{\sc pos} taggers to chunk parsers and intermediate parsers such as
{\sc rasp} \cite{briscoe:etal:2006}.  {\sc mrs} analyses derived from
deep grammars, such as the English Resource Grammar ({\sc erg},
\cite{copestake:flickinger:2000}) can be seen as special cases of {\sc
  rmrs}.

However, from a formal perspective, \rmrs\ is somewhat ad-hoc in that
it has no clearly defined model theory; i.e., it is not defined what,
exactly, any given \rmrs\ representation actually \emph{means}.  The
key contribution we make is to cast {\sc rmrs}, for the first time, as
a logic with a well-defined model theory.  This has a number of
advantages.  First, it allows us to formally relate \rmrs\ to other
frameworks such as \mrs\ and predicate logic in order to establish the
compatibility of \rmrs\ with the approaches to robust semantic
analysis mentioned above.  \todo{except we don't do this here --
  perhaps we should} Second, it provides us with a notion of logical
entailment, which we can use to verify the compatibility of \rmrs\
representations computed from different parsers, such as a deep and a
shallow parser.  Finally, it is only through rigid formalisation that
we can give users of \rmrs\ access to inference systems and to
algorithms for efficiently computing the possible fully specific
semantic representations described by the \rmrs\ output of a shallow
parser.

>>>>

On the other hand, there has been a recent trend towards systems for
robust wide-coverage semantic construction.  For instance,
\newcite{bos:etal:2004} exploit {\sc ccg}'s syntax-semantics
interface to produce a unique logical form for each derivation of a
statistical {\sc ccg} parser; \newcite{zettlemoyer:collins:2007} and
others learn string-to-logical-form mappings, exploting informative
features from shallow language processors; and
\newcite{copestake:2003} and \newcite{frank:2004} endow robust
language processors with semantic components.

@@

There are certain semantic phenomena that these robust
approaches don't capture reliably, including quantifier scope,
optional arguments, and long-distance dependencies (for instance,
\newcite{clark:etal:2004} report that the parser used by
\newcite{bos:etal:2004} yields only 63\% accuracy on object
extraction; e.g., {\em the man that I met\dots}).  Forcing a shallow
parser to make a decision about these phenomena can therefore be
error-prone.  Depending on the application, it may be preferrable to
give the parser the option to leave the decision open when it's not
sufficiently informed---i.e., to compute a partial semantic
representation and to complete it later, using information
extraneous to the parser.

In this paper, we focus on one particular approach to representing
partial semantic information: Robust Minimal Recursion Semantics
(\rmrs, \cite{copestake:2003}).  \rmrs\ is a generalisation of \mrs\
\cite{copestake:etal:2005} that is designed to support
underspecification of scope and predicate-argument structure.  It is
an emerging standard representation language for partial semantic
information, and has been applied in several implemented systems.  For
instance, \newcite{copestake:2003} and \newcite{frank:2004} use it to
specify semantic components to shallow parsers ranging in depth from
{\sc pos} taggers to chunk parsers and intermediate parsers such as
{\sc rasp} \cite{briscoe:etal:2006}.  {\sc mrs} analyses derived from
deep grammars, such as the English Resource Grammar ({\sc erg},
\cite{copestake:flickinger:2000}) are special cases of {\sc
  rmrs}.

However, from a formal perspective, \rmrs\ is somewhat ad-hoc in that
it has no clearly defined model theory; i.e., what exactly, any given
\rmrs\ representation actually \emph{means} is not defined.  The key
contribution we make is to cast {\sc rmrs}, for the first time, as a
logic with a well-defined model theory.  This has a number of
advantages.  First, it allows us to formally relate \rmrs\ to other
frameworks such as \mrs\ and predicate logic in order to establish the
compatibility of \rmrs\ with the approaches to robust semantic
analysis mentioned above.  
Second, it provides us with a notion of logical
entailment, which we can use to verify the compatibility of \rmrs\ 
representations computed from different parsers, such as a deep and a
shallow parser.  Finally, it is only through rigid formalisation that
we can give users of \rmrs\ access to inference systems and to
algorithms for efficiently computing the possible fully specific
semantic representations described by the \rmrs\ output of a shallow
parser.

### 2008-09-18 trunk/Papers/rmrs-08/motivation.tex
18/09/2008

We start with an informal discussion of some toy \rmrs s, which are
meant to illustrate how \rmrs\ can be used to represent partial
semantic information, and why people find it useful for this purpose.

Consider the following sentence:
\begin{examples}
  \item Every fat cat chased some dog.
\end{examples}

@@

\begin{examples}
\item $\sem{\_every\_q\_1}(x, \sem{\_fat\_j\_1}(e',x) \wedge
    \sem{\_cat\_n\_1}(x), \sem{\_some\_q\_1}(y, \sem{\_dog\_n\_1}(y),
    \sem{\_chased\_v\_1}(e,x,y)))$
\label{ex:fat-cat-1}
\item $\sem{\_some\_q\_1}(y, \sem{\_dog\_n\_2}(y),
\sem{\_every\_q\_1}(x, \sem{\_fat\_j\_1}(e',x) \wedge
    \sem{\_cat\_n\_2}(x), 
    \sem{\_chased\_v\_1}(e,x,y)))$
\label{ex:fat-cat-2}
\end{examples}

@@

Now imagine that we are trying to extract semantic information from
the output of a part-of-speech tagger by using the word lemmas as
lexical predicate symbols.  Such a semantic representation is highly
partial, as it will not say anything about predicate-argument
relations, or resolve lexical ambiguity -- it will use predicate
symbols such as $\sempred{\_cat\_n}$, which might resolve to the
predicate symbols $\sem{\_cat\_n\_1}$ or $\sem{\_cat\_n\_2}$ in the
complete semantic representation.  (Notice that we are using different
fonts for the ambiguous and unambiguous predicate symbols.)  We can
express this information in \rmrs\ as follows.

\begin{examples}
\item \label{ex:cat-pos}
$l_1:a_1:\sempred{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\sempred{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\sempred{\_cat\_n}(x_3)$\\
$l_5:a_5:\sempred{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\sempred{\_some\_q}(x_4)$, \\
$l_9:a_9:\sempred{\_dog\_n}(x_5)$
\end{examples}

This \rmrs\ consists of six \emph{elementary predications} (EPs), one
for each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors into each tree in such a way that the label of
the node they are mapped to is consistent with the label specified in
the elementary predication.  For instance, the first tree satisfies
the \rmrs\ because we can map $l_1$ and $a_1$ to the root of that
tree, and the root label $\sem{\_every\_q\_1}$ is consistent with the
EP predicat $\sempred{\_every\_q}$.

There are of course many other trees (and thus, fully specific
semantic representations such as (\ref{ex:fat-cat-1})) that are
described equally well by the same \rmrs; this is not surprising,
given that the semantic information we got from the POS tagger is so
incomplete.  If we have more information, say information about the
predicate-argument structure and semantic dependencies from a partial
parser such as \todo{what? CASS?}, we can represent these in a more
detailed \rmrs, as follows:

\begin{examples}
\item 
$l_1:a_1:\sempred{\_every\_q}(x)$, \\
$l_{41}:a_{41}:\sempred{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\sempred{\_cat\_n}(x_3)$\\
$l_5:a_5:\sempred{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\sempred{\_some\_q}(x_4)$, \\
$l_9:a_9:\sempred{\_dog\_n}(x_5)$
\todo{add more stuff with ARG and equalities here}
\label{ex:cat-partial-parser}
\end{examples}

This \rmrs\ uses two new types of atoms.  Atoms of the form $x=y$
express that the two variables $x$ and $y$ should really be the same;
in the example, the \rmrs\ uses two variable names $x$ and $x'$, which
are mapped to the same variable $x$ in Fig.~\ref{fig:1}.  \todo{I'm
  making these names up -- adjust to RMRS.}  Atoms of the form
$\ARG_i(a,x)$ and $\ARG_i(a,h)$ express that a node representing the
variable $x$ or a node denoted by the \emph{hole} $h$ is the $i$-th
child of the node to which the anchor $a$ refers.  The separation of
EPs and $\ARG$ atoms is perhaps the most important difference between
\rmrs\ and earlier underspecification formalisms such as \mrs\ and
dominance graphs, which required that a predicate and its arguments
had to be specified together.  \rmrs\ goes beyond these formalisms in
expressive power, especially if we allow atoms of the form
$\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether $x$ is the
second or third child of the anchor $a$.

Notice also that the labels of the EPs for ``fat'' and ``cat'' are
stipulated to be equal in the \rmrs, whereas the anchors are not.
In the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_a\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an EP is to carry the predicate and connect to
the arguments; the role of the label is to connect the EP to the
surrounding formula.  The use of equal labels for two EPs stems from
\mrs, where intersective modification was also modeled by using two
EPs with the same label.

@@

\begin{examples}
\item $l_1:a_1\handel\mbox{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mbox{RSTR}(a_1,h_2),
\mbox{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\mbox{\_fat\_j\_1}(e'), \mbox{ARG1}(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\mbox{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\mbox{\_chase\_v\_1}(e_{spast})$,\\
\hspace*{0.1in}$\mbox{ARG1}(a_5,x_4),
\mbox{ARG2}(a_5,x_5)$\\ 
$l_6:a_6\handel\mbox{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mbox{RSTR}(a_6,h_7),
\mbox{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\mbox{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\todo{fix RSTR and BODY to ARG1 and ARG2? or talk about it in the text?}
\label{ex:cat-erg}
\end{examples}

This is now an extremely detailed underspecified representation of the
different semantic representations of the sentence, which is
essentially a notational variant of the \mrs\ that the ERG would
derive.  Notice the use of atoms like $h_2 \qeq l_{42}$, which specify
a kind of ``outscopes'' relationship between the hole and the label,
and can be used to underspecify the scope of the two quantifiers.

>>>>

We start with an informal discussion to 
illustrate how \rmrs\ can be used to represent partial
semantic information, and why people find it useful for this purpose.
Consider the following (toy) sentence:
\begin{examples}
  \item Every fat cat chased some dog.
\end{examples}

@@

\begin{examples}
\item $\sem{\_every\_q\_1}(x, \sem{\_fat\_j\_1}(e',x) \wedge
    \sem{\_cat\_n\_1}(x),$\\
\hspace*{0.1in} $\sem{\_some\_q\_1}(y, \sem{\_dog\_n\_1}(y),$\\
\hspace*{0.2in}$\sem{\_chased\_v\_1}(e,x,y)))$
\label{ex:fat-cat-1}
\item $\sem{\_some\_q\_1}(y, \sem{\_dog\_n\_2}(y),$\\
\hspace*{0.1in}$\sem{\_every\_q\_1}(x, \sem{\_fat\_j\_1}(e',x) \wedge
    \sem{\_cat\_n\_2}(x), $\\
\hspace*{0.2in}$\sem{\_chased\_v\_1}(e,x,y)))$
\label{ex:fat-cat-2}
\end{examples}

@@

Now imagine that we are trying to extract semantic information from
the output of a part-of-speech tagger by using the word lemmas as
lexical predicate symbols.  Such a semantic representation is highly
partial, as it will not say anything about predicate-argument
relations or resolve lexical ambiguity -- it will use predicate
symbols such as $\sempred{\_cat\_n}$, which might resolve to the
predicate symbols $\sem{\_cat\_n\_1}$ or $\sem{\_cat\_n\_2}$ in the
complete semantic representation.  (Notice that we are using different
fonts for the ambiguous and unambiguous predicate symbols.)  
We can
express this information in \rmrs\ as follows.

\begin{examples}
\item \label{ex:cat-pos}
$l_1:a_1:\sempred{\_every\_q}(x_1)$, \\
$l_{41}:a_{41}:\sempred{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\sempred{\_cat\_n}(x_3)$\\
$l_5:a_5:\sempred{\_chased\_v}(e_{past})$, \\
$l_6:a_6:\sempred{\_some\_q}(x_6)$, \\
$l_9:a_9:\sempred{\_dog\_n}(x_7)$
\end{examples}

This \rmrs\ consists of six \emph{elementary predications} ({\sc ep}s), one
for each word lemma in the sentence; each of them is prefixed by a
\emph{label} and an \emph{anchor}, which are essentially variables
that refer to nodes in the trees in Fig.~\ref{fig:1}.  
The lack of semantic dependencies rendered by a {\sc pos} tagger
also means we have unique arguments to each
predicate symbol (e.g., see $x_1$ and $x_3$ in (\ref{ex:cat-pos}),
compared with the {\em same} variable $x$ in the semantic
representations (\ref{ex:fat-cat-1}) and (\ref{ex:fat-cat-2})).
We can say that
the two trees \emph{satisfy} this \rmrs\ because it is possible to map
the labels and anchors into nodes in each tree and variable names like
$x_1$ and $x_3$ into variable names in the tree in such a way that the
predication
of
the node that labels and anchors are mapped to is consistent with the
that given in the {\sc ep}s of the {\sc rmrs}.
For instance, the first tree satisfies
the \rmrs\ because we can map $l_1$ and $a_1$ to the root of that
tree, $x_1$ (and $x_3$) are both mapped to $x$, and the root label
$\sem{\_every\_q\_1}$ is consistent with the 
{\sc ep} predicate $\sempred{\_every\_q}$.

There are of course many other trees (and thus, fully specific
semantic representations such as (\ref{ex:fat-cat-1})) that are
described equally well by the same \rmrs; this is not surprising,
given that the semantic information we got from the POS tagger is so
incomplete.  If we have more information, say information about
subject and object relations from a chunk parser like Cass
\cite{abney:1996}, we can represent these in a more
detailed \rmrs, as follows:

\begin{examples}
\item 
$l_1:a_1:\sempred{\_every\_q}(x_1)$, \\
$l_{41}:a_{41}:\sempred{\_fat\_a}(e')$,\\
$l_{42}:a_{42}:\sempred{\_cat\_n}(x_3)$\\
$l_5:a_5:\sempred{\_chased\_v}(e_{past})$, 
\hspace*{0.1in} $\ARG1(a_5,x_4),
\ARG2(a_5,x_5)$\\ 
$l_6:a_6:\sempred{\_some\_q}(x_6)$, \\
$l_9:a_9:\sempred{\_dog\_n}(x_7)$\\
$x_3=x_4$, $x_5=x_7$
\label{ex:cat-partial-parser}
\end{examples}

This \rmrs\ uses two new types of atoms.  Atoms of the form $x_3=x_4$
express that the two variables $x_3$ and $x_4$ must map to the same
variable in any fully specific logical form; e.g., to the variable $x$
in Fig.~\ref{fig:1}.  Atoms of the form
$\ARG_i(a,x)$ and $\ARG_i(a,h)$ express that the $i$-th child of the
node to which the anchor $a$ refers is the variable name that $x$ maps
to (or the node that the {\em hole} $h$ denotes).  The separation of
{\sc ep}s and $\ARG$ atoms is perhaps the most important difference between
\rmrs\ and earlier underspecification formalisms such as \mrs\ and
dominance graphs, which required that a predicate and its arguments
had to be specified together.  \rmrs\ goes beyond these formalisms in
expressive power, because it allows one to state separately
information about how many arguments a predicate takes, and the values of
those (non-scopal and scopal) arguments.   If we also allow atoms of the form
$\ARG_{\{2,3\}}(a,x)$ to express uncertainty as to whether $x$ is the
second or third child of the anchor $a$, then \rmrs\ can specify the
arguments to a predicate while underspecifying their position.

@@

\begin{examples}
\item $l_1:a_1\handel\mbox{\_every\_q\_1}(x_1),$\\
\hspace*{0.1in}$\mbox{RSTR}(a_1,h_2),
\mbox{BODY}(a_1,h_3)$\\ 
$l_{41}:a_{41}\handel\mbox{\_fat\_j\_1}(e'), \ARG1(a_{41},x_2)$\\
$l_{42}:a_{42}\handel\mbox{\_cat\_n\_1}(x_3)$\\
$l_5:a_5\handel\mbox{\_chase\_v\_1}(e_{spast})$,\\
\hspace*{0.1in}$\ARG1(a_5,x_4),
\ARG2(a_5,x_5)$\\ 
$l_6:a_6\handel\mbox{\_some\_q\_1}(x_6)$,\\
\hspace*{0.1in}$\mbox{RSTR}(a_6,h_7),
\mbox{BODY}(a_6,h_8)$\\ 
$l_9:a_9\handel\mbox{\_dog\_n\_1}(x_7)$\\
$h_2=_q l_{42}, l_{41}=l_{42}, h_7 =_q l_9$\\
$x_1=x_2, x_2=x_3, x_3=x_4,$\\
$x_5=x_6, x_5=x_7$
\label{ex:cat-erg}
\end{examples}
RSTR and BODY are conventional names for the scopal arguments to
predicate symbols corresponding to quantifiers.  
This is now an extremely detailed underspecified representation of the
different semantic representations of the sentence, which is
essentially a notational variant of the \mrs\ that the {\sc erg} would
derive.  Notice the use of atoms like $h_2 \qeq l_{42}$, which specify
a kind of ``outscopes'' relationship between the hole and the label,
and can be used to underspecify the scope of the two quantifiers.

Notice also that the labels of the EPs for ``fat'' and ``cat'' are
stipulated to be equal in the \rmrs\ (\ref{ex:cat-erg}), whereas the
anchors are not. 
In the tree, it is the anchors that are mapped to the nodes with the
labels $\sem{\_fat\_a\_1}$ and $\sem{\_cat\_n\_1}$; the label is
mapped to the conjunction node just above them.  In other words, the
role of the anchor in an {\sc ep} is to connect a predicate to its
arguments, while the role of the label is to connect the {\sc ep} to the
surrounding formula.  Representing conjunction with label sharing
stems from \mrs\ and provides compact semantic representations.
